Wikipedia Text Mining and Surprisal Analysis

### 2019-07-10 surprisal/WALS_features.Rmd
using existing features data

```{r setup, include=FALSE}
library(tidyverse)
library(lingtypology)
library(feather)
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

@@

FEATURES <- c(1:144)
for (i in c(1:length(FEATURES))) {
  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
} #get all word order WALS features in correct syntax

wf <- wals.feature(FEATURES)
```

```{r see which WALS features are the most complete for our languages}
df <- wf %>% 
  filter(language %in% LANGUAGES) %>% 
  filter(rowSums(is.na(.)) < 130) %>% 
  select(`30A`:`121A`) %>% 
  is.na() %>% 
  colSums() %>% 
  sort(decreasing = T)
```

```{r get common languages between the imputations and models}
nom_categories <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `30A`:`57A`) %>%
  filter(rowSums(is.na(.)) < 12)

nom_syntax <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `58A`:`64A`) %>%
  filter(rowSums(is.na(.)) < 5)

verb_categories <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `65A`:`80A`) %>%
  filter(rowSums(is.na(.)) < 8)

word_order <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `81A`:`97A`) %>%
  select(-`84A`)

clauses <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `98A`:`121A`) %>%
  filter(rowSums(is.na(.)) < 12)

common_df <- nom_categories %>% 
  semi_join(nom_syntax) %>%
  semi_join(verb_categories) %>%
  semi_join(word_order) %>%
  semi_join(clauses)

@@

```{r comparison: features to cosines}
imputed_df <- common_df %>% 
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = factor(value)) %>%
  select(language, feature, value) %>%
  ungroup() %>% 
  spread(feature, value) %>% 
  sapply(as.factor) %>%
  as.data.frame() %>%
  column_to_rownames("language") %>%
  MIMCA(ncp = 1)

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(lingtypology)
library(feather)
library(here)
library(missMDA)
knitr::opts_chunk$set(echo = TRUE)
```

@@

wals <- read.csv(here("Data/wals.csv"))
feature_names <- read_feather(here("Wikipedia/feature_names.feather"))
```

```{r see which WALS features are the most complete for our languages}
all_languages <- wals %>% 
  filter(language %in% LANGUAGES) %>% 
  filter(rowSums(is.na(.)) < 130) %>%
  group_by(language) %>%
  slice(1) %>%
  ungroup() 
```

```{r get common languages between the imputations and models}
nom_categories <- all_languages %>%
  select(language, X30A:X57A) %>%
  filter(rowSums(is.na(.)) < 12)

nom_syntax <- all_languages %>%
  select(language, X58A:X64A) %>%
  filter(rowSums(is.na(.)) < 5)

verb_categories <- all_languages %>%
  select(language, X65A:X80A) %>%
  filter(rowSums(is.na(.)) < 8)

word_order <- all_languages %>%
  select(language, X81A:X97A) %>%
  select(-X84A)

clauses <- all_languages %>%
  select(language, X98A:X121A) %>%
  filter(rowSums(is.na(.)) < 12)

common_df <- nom_categories %>% 
  inner_join(nom_syntax) %>%
  inner_join(verb_categories) %>%
  inner_join(word_order) %>%
  inner_join(clauses)

@@

```{r comparison: features to cosines}
#imputed_df <- 
  common_df %>% 
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = factor(value)) %>%
  select(language, feature, value) %>%
  ungroup() %>% 
  spread(feature, value) %>% 
  sapply(as.factor) %>%
  as.data.frame() %>%
  column_to_rownames("language") %>% pull(X98A) %>% length()
  MIMCA(ncp = 1)

### 2019-07-10 surprisal/wiki_surprisal.Rmd
using existing features data

Always run this section. It takes time, but this gets all of the WALS features for every language in WALS along with the unigram slope data for all of the Wikipedia languages we've run.

```{r get features, eval=F}
wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))
LANGUAGES <- wkpd$language

FEATURES <- c(1:144)
for (i in c(1:length(FEATURES))) {
  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
} #get all word order WALS features in correct syntax

wf <- wals.feature(FEATURES)

write_csv(wf, here("Data/wals.csv"))
```

```{r load-wals, eval = T}
wf <- read_csv(here("Data/wals.csv"))
```

@@

```{r nominal categories}
wo_wf <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `30A`:`57A`) %>%
  filter(rowSums(is.na(.)) < 12)

```
```{r nominal syntax}
wo_wf <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `58A`:`64A`) %>%
  filter(rowSums(is.na(.)) < 5)

```
```{r verbal categories}
wo_wf <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `65A`:`80A`) %>%
  filter(rowSums(is.na(.)) < 8)

```
```{r word order}
wo_wf <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `81A`:`97A`) %>%
  select(-`84A`)
```
```{r clauses}
wo_wf <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `98A`:`121A`) %>%
  filter(rowSums(is.na(.)) < 12)

>>>>

Always run this section. This gets all of the WALS features, for every language, along with the unigram slope data for all of the Wikipedia languages we've run.

```{r get features, eval=T}
wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))
LANGUAGES <- wkpd$language

wf <- read_csv(here("Data/wals.csv"))

@@

```{r nominal categories, eval=F}
wo_wf <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, X30A:X57A) %>%
  filter(rowSums(is.na(.)) < 12)

```
```{r nominal syntax, eval=F}
wo_wf <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, X58A:X64A) %>%
  filter(rowSums(is.na(.)) < 5)

```
```{r verbal categories, eval=F}
wo_wf <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, X65A:X80A) %>%
  filter(rowSums(is.na(.)) < 8)

```
```{r word order, eval=F}
wo_wf <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, X81A:X97A) %>%
  select(-`84A`)
```
```{r clauses, eval=F}
wo_wf <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, X98A:X121A) %>%
  filter(rowSums(is.na(.)) < 12)
